---
title: "Class15week7thurs"
author: "Emily Kimball"
date: "11/14/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##import count and metadata 
```{r}
metafile <- "GSE37704_metadata.csv"
countfile <- "GSE37704_featurecounts.csv"
```
#import meta data and take a peak
```{r}
colData = read.csv(metafile, row.names=1)
head(colData)

countData = read.csv(countfile, row.names=1)
head(countData)
```
#remove the column 
```{r}
countData <- as.matrix(countData[, -1])
head(countData)
```
#matching up the data making sure that the things are lined up correctly
```{r}
colnames(countData)
rownames(colData)
#this shows that all of the values are matching up with eachother
all(colnames(countData) == rownames(colData))
```

#Filter out the data that has 0 as a result so remove any data with zeros 
```{r}

countData = countData[ rowSums(countData) > 0,]
head(countData)


```
Load the data library 
```{r}
library(DESeq2)
```

```{r}
#Setup the object with our data in the way DEseq wants it 

dds = DESeqDataSetFromMatrix(countData=countData,
                             colData=colData,
                             design=~condition)
#Run the analysis 
dds = DESeq(dds)
```
#getting the results 
```{r}
res = results(dds)
res
```
#plot the results 
```{r}
plot(res$log2FoldChange, -log(res$padj))

#making a color vector 
mycols <- rep("grey", nrow(res))
#Making red color 
"red" <- mycols[abs(res$log2FoldChange > 2)]
#making blue for p value less than 0.01 
inds <- mycols[abs(res$padj < 0.01)] & mycols[abs(res$log2FoldChange) > 2]
"blue" <- mycols[inds]

plot(res$log2FoldChange, -log(res$padj), col=, xlab="Log2(FoldChange)", ylab="-Log(P-value)" ))

```
#gene annotation 
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

#we can translate between all of the following database ID systems 
columns(org.Hs.eg.db)

res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(countData),     #where are my IDs
                    keytype="ENSEMBL",              #what format are my IDs
                    column="SYMBOL",                # the new format I want 
                    multiVals="first")

res$entrez = mapIds(org.Hs.eg.db,
                    keys= row.names(countData),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

res$name =   mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="GENENAME",
                    multiVals="first")

head(res, 10)
```
##Pathway analysis
Here we are going to use the gage package and pathway pachage to draw pathway diagrams 
```{r}
library(pathview)
library(gage)
library(gageData)

data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
```
Making a litte vector
```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)

# Get the results in the gage pathway
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```
#look at the results returned from gage

```{r}
attributes(keggres)

#heading the data 
head(keggres$greater)
head(keggres$less)
```

#Now trying out the pathway package 
```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```
A different way to get a result 
```{r}
# A different PDF based output of the same data
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```
#focusing on frewer pathways 
```{r}
## Focus on top 5 upregulated pathways here for demo purposes only
keggrespathways <- rownames(keggres$greater)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids

#this function makes the graphs 
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")
```
#Doing the same thing for the downregulated pathways 
```{r}
downreg <- rownames(keggres$less)[1:5]
keggresids = substr(downreg, start = 1, stop = 8)
pathview(gene.data = foldchanges, pathway.id = keggresids, species = "hsa")
```

#Doing the same thing but for gene ontology 
```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```

#Part 4:
















