---
title: "CLass 11: Structural Bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database for biomolecular structures
>Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can
you determine what proportion of structures are protein? Aim to have a rendered GitHub
document with working code that yields your answers.

```{r}
read.csv("Data Export Summary.csv")
data <- read.csv("Data Export Summary.csv")
#answer the question 
#Total entries 
 total <- sum(data$Total)
 #% xray vs mircroscope
data$Total/total *100

##proportion protien 
round( data$Proteins/(sum(data$Proteins))* 100, 2 )

```
```{r}
#section 4.2 on the inclass lab 
library(bio3d)
 hsg <- read.pdb("1hsg (1).pdb")
whsg <- write.pdb(read.pdb("1hsg (1).pdb"))
 ca.inds <- atom.select(hsg, "calpha")
#print details of the first selected atoms 
head( hsg$atom[ca.inds$atom, ] )
```

```{r}
# Select chain A
a.inds <- atom.select(hsg, chain="A")
# Select C-alphas of chain A
ca.inds <- atom.select(hsg, "calpha", chain="A")
# We can combine multiple selection criteria to return their
#intersection
cab.inds <- atom.select(hsg, elety=c("CA","CB"), chain="A",
resno=10:20)
```



> showing how you can select different aspects 

```{r}
ligand <- atom.select( hsg, "ligand", value = TRUE)
write.pdb(ligand, file = "1hsg (1).pdb")
```


