---
title: "Class 14 Week 7 Tuesday"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Reading the Data 
```{r}
metadata <- read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
counts <-  read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
```
#Counting the Metadata 
```{r}
head(counts)
head(metadata)
```
#How many exps 
```{r}
#ncol(counts)
#how many genes 
#nrows(counts)
```

#Getting a table with just the control 
```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/4 
names(control.mean) <- counts$ensgene
```
#Getting the treated for meta data 
```{r}
treated <- metadata[metadata[, "dex"]== "treated",]
treated.mean <- rowSums( counts[ , treated$id])/4
names(treated.mean) <- counts$ensgene
```

#getting the control means for both 
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```
#colsums the data 
```{r}
colSums(meancounts)
```
#make a dataplot 
we want to know if there is a difference in expression values for control vs treatement.  
The first step is to find which experiments were the control and which were the treated and then get the average values across all control and treated sides.  

```{r}
plot(meancounts$control, meancounts$treated, log = "xy")

```
#need to scale the data by log2 to make it more readable 
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
#We need to account for the funky values Nan and -inf this code will show all the values even if they have a level of expression close to 0

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind = TRUE)
to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```
#showing which things are actually significant by using 2foldchanges 
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

#count how many up and down regulated genes there are 
```{r}
 upreg <- sum(up.ind)
 downreg <- sum(down.ind)
```
#adding another file into our annotation 
```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```
#loading the DESeq2 dataset and creating the dds
```{r}
library(DESeq2)
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```
#Running the packages 
```{r}
sizeFactors(dds)
dispersions(dds)

dds <- DESeq(dds)
```
#getting the results 
```{r}
#results 
res <- results(dds)
res
#summary of the res 
summary(res)
#need to reorder the values 
resOrdered <- res[order(res$pvalue),]
#setting a new value for the pvalue 
res05 <- results(dds, alpha=0.05)
summary(res05)
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)
```
#making a volcano plot 
##adding all the ablines so we can then color it 
```{r}
plot(res$log2FoldChange, -log(res$padj))
abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)

#setting up a color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ] <- "red"
inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

#Volcano plot with custom colors
plot( res$log2FoldChange, -log(res$padj),
col=mycols, ylab="-Log(P-value)",
xlab="Log2(FoldChange)" )
#saving our results 
write.csv(res, file = "expression_results.csv")
```




