---
title: "class7week4tues"
author: "Emily Kimball"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## R functions revisited 
```{r}
source("http://tinyurl.com/rescale-R")
```


```{r}
rescale(1:10)
```
the function is already correct, becase we fixed them in the previous class
```{r}
rescale(c(1,10,5,NA,6))
```

practicing on rescale2  -> this is the result we want 
```{r}
rescale2(c(1:10, 3))
```


##more examples with functions 
```{r}
x <- c(1,2,NA,3,NA)
y <- c(NA, 3, NA, 3, 4)

#google for a function that might help to find where the NA elements in 2 input vectors 
is.na(x)
```
```{r}
which(is.na(x))
```

this gives us a value that combines our function and only gives one result
THIS is my working snippet!!!
```{r}
sum(is.na(x) & is.na(y))
```
this shows that the sum will treat TRUE and FALSE as 1, 0
```{r}
sum(c(TRUE, FALSE, TRUE, TRUE, FALSE))
```
now turning this into a function
```{r}
both_na <- function(x, y) {sum(is.na(x) & is.na(y))}
```
testing out the functiton 
```{r}
both_na(x,y)
```
#idiot proofing the silly mistakes 
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
#testing in out
both_na(x,y2)
```
Writing a function to heck 
```{r}
both_na2 <- function(x, y) {if(length(x) != length(y)) {
  stop("vectors should be the same length")
}
  sum(is.na(x) & is.na(y))
  }
```
Testing out the function 
```{r}
both_na2(x, y2)
```
#do it yourself 
```{r}
# student 1
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
# student 2
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80) 
 #this is our snippet 
mean(student1[-which.min(student1)], na.rm = TRUE)
mean(student2[-which.min(student2)], na.rm = TRUE) 

```
We have our working code to turn into a function
```{r}
grade <- function(x) {
   if(any(is.na(x))){
       warning("Student is missing a homework grade")}
  mean(x[-which.min(x)], na.rm = TRUE)
 }
print(grade)
```
```{r}
grade(student1)
grade(student2)
```
#Using the class data 
```{r}
url <-  "https://tinyurl.com/gradeinput"
hw <- read.csv(url, row.names=1)
```
using apply
```{r}
apply(hw,1,grade)

```

#Installing bioconductor 
```{r}
install("dplyr")
```




