---
title: "Lect9week5tues"
author: "Emily Kimball"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#read.csv("WisconsinCancer.csv")
# Save your input data file to a new 'data' directory

# Complete the following code to input the data and store as wisc.df
wisc.df=read.csv("WisconsinCancer.csv")
head(wisc.df)
```

```{r}
wisc.data <- as.matrix(wisc.df[,3:32])

#naming rows 
row.names(wisc.data) <- wisc.df$id

head(wisc.data)

diagnosis <- wisc.df$diagnosis

```
#making the disgnosis how many are maligement 
```{r}
table(wisc.df$diagnosis)
```
Q3. How many variables are suffixed with mean
Grep with will group together (pattern, x, ...)
```{r}
grep("_mean",colnames(wisc.df))
#how many --> we can use legnth to count the # of matches 
length(grep("_mean",colnames(wisc.df)))

```
##PCA analysis 
```{r}
#check column means 
colMeans(wisc.data)
apply(wisc.data, 2, sd)
```

```{r}
# we must set scale to true, becuase the data is affected by sigfigs
wisc.comp <- prcomp(wisc.data, scale = TRUE)

summary(wisc.comp)
biplot(wisc.comp, scale = TRUE)

#this plot is bad so we will try to make a dot plot 
# Scatter plot observations by components 1 and 2

plot(wisc.comp , col =  , 
     xlab = "PC1", ylab = "PC2")
```
Making a Plot of PC1 and Pc2 based upon maligment or benign 
```{r}
plot(wisc.comp$x[,1], wisc.comp$x[,2], col=diagnosis)
diagnosis
```

```{r}
diagnosis
```
#varience explained 
```{r}
pr.var <- (wisc.comp)
head(pr.var)
```

#Hierachal clustering
```{r}
 data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "complete")


#making the plot 
plot(wisc.hclust)
abline(, col="red", lty=2)
```
#that is a hard graph to read so lets try again 
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, 4)
table(wisc.hclust.clusters, diagnosis)
```

```{r}
wisc.km <- kmeans(wisc.hclust.clusters, centers= 2 , nstart= 20)
table(wisc.km$cluster, diagnosis)
```
##Combining Methods Doing the new method 
```{r}
 newdata <- dist(wisc.comp$x[,1:7])
hclusdata <- hclust(newdata, method ="ward.D2")
plot(hclusdata)

grps <- cutree(hclusdata, k = 2)
table(grps)
table(grps, diagnosis)
```
#Plotting this data to get a graph coloered by groups  
```{r}
plot(wisc.comp$x[,1:2], col=grps)
```
#reorganizing the data 
```{r}
g <- as.factor(grps)
#relevel 
g <- relevel(g, 2)
#now we plot based on the new variable color 
plot(wisc.comp$x[,1:2], col=g)
```
#Predictions 
```{r}
 new<- read.csv("new_samples.csv")
```
```{r}
npc <- predict(wisc.comp, newdata=new)
npc
```
#graphing the new data against the old 
```{r}
plot(wisc.comp$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
#The data on the grapph below is the predicted vs actual 
```
based upon the patients we should look at patient one more seriously 





